---
title: "ML-lab2-Bakukha"
author: "Bakukha-IAB21"
date: "13 05 2022"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Read the data
```{r}
data <- read.csv("mlbootcamp5_train.csv", sep=";")
head(data)
```

Transform the features: 
- create "age in years" (`age_years`)dividing age by 365.25 and taking floor ($\lfloor{x}\rfloor$ is the largest integer that is less than or equal to $x$)
- remove "age" feature
- transfrom `cholesterol` and `gluc` to factor

```{r}
data$age_years <- floor(data$age / 365.25)
data <- subset(data, select=-c(age))
data$cholesterol <- as.factor(data$cholesterol)
data$gluc <- as.factor(data$gluc)
data$cardio <- as.factor(data$cardio)
head(data)
```

Split data into train and test parts in the proportion of 7/3.
The target feature is `cardio`

```{r}
library(caret)
part <- createDataPartition(data$cardio, p=0.7, list=FALSE, times=1)
train_data <- data[part,]
test_data <- data[-part,]
head(train_data)
head(test_data)
```

Train a decision tree on the dataset `training` with **max depth equal to 3**. For setting this parameter use function `rpart::rpart.control`. Use default cross calidation parameters.

```{r}
library(rpart)
library(caret)
model <- train(cardio ~ ., data = train_data, method = "rpart", control = rpart.control(maxdepth = 3))

model
```

Plot final tree with `rattle` library

```{r}
library(rattle)
fancyRpartPlot(model$finalModel, main="Tree plot")
```

What is the accuracy of prediction on test data?

```{r}
prediction <- predict(model, test_data)
accuracy <- sum(prediction == test_data$cardio) / nrow(test_data)
accuracy
```
Now set cross validation parameters: `method = "cv"`, `number = 5`. Tree parameters set to default.

```{r}
trControl <- trainControl(method = "cv", number = 5)
model2 <- train(cardio ~ ., data = train_data, method = "rpart", trControl = trControl)
model2
```

Plot final tree with `rattle` library

```{r}
library(rattle)
fancyRpartPlot(model2$finalModel, main="Tree plot")
```

What is the accuracy of prediction on test data?

```{r}
prediction2 <- predict(model2, test_data)
accuracy2 <- sum(prediction2 == test_data$cardio) / nrow(test_data)
accuracy2
```
Does the accuracy became better or not?

**Accuracy is the same**